Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2024-08-06 doc/manuscript/response-to-reviews.Rmd

The aims of the study are clearly identified and very relevant to the community, and the idea itself of improving scientist’s contribution to international databases appreciable. Thus, there are good reasons to publish.

**Thank you.**

The results section is however disconcerting, partly because it is not clearly structured, and partly because it is quite unclear what point authors wanted to make out of the few graphs presented.

**To help clarify the message, we have responded to specific concerns (detailed below) and done some reorganization and rewording to make it less verbose. Our aim has been to present an objective overview of the ForC data contributed to EFDB and the potential for future submissions to improve EFDB data volume and composition.**

@@

The first paragraph is about records with deficiencies, a fraction being “deemed not applicable to EFDB…”. This is not informative at all in absence of a series of well-identified examples. Quite verbose but not that useful.

**We have added a sentence detailing some of the reasons for rejection: "Reasons included -- but were not limited to -- non-applicability to the IPCC methodology of the variable submitted (e.g., net ecosystem CO~2~ exchange, litter - OL layer; subsequently excluded from lists of relevant variables and counts of relevant records), inadequate information on the quality of data and thus on its replicability (e.g., confidence interval/uncertainty), and vaguely described disturbance histories (e.g., "moderately"/ "severely" burned)." We also note that the authors of this study were not privy to the discussions of the review panel and do not have detailed information on all reasons for rejection of each record.**

**In addition, we have reworded and reorganized this paragraph (and the following) to make it less verbose.**

@@

Another example, the map (Fig 2) is fine as an illustration or proof of the global participation to the efforts of reporting, but it doesn’t help identify the gaps in the current reporting. Some vegetation types are well represented, others have a serious deficit. Identifying and quantifying this, for instance in terms of report per area, would have been a more eloquent and useful than a map. The imbalance in the data goes for the below/aboveground compartments too.

<!--**We have made many of the suggested revisions (detailed below) to improve the assessment of data representativeness and consider this to be an important improvement. At the same time, we also wish to emphasize that the goal of this paper was to document the process and provide an overview of the types of data submitted, as opposed to submitting data that adequately represents all forest types and documenting their representativeness. The latter -- to the extent that it would even be possible -- would be a massive project far beyond the scope of this paper. Therefore, while improving the assessment of data representativeness, we believe it would be overkill to analyze all aspects of representativeness.**-->

### 2024-08-06 doc/manuscript/ForC-EFDB manuscript.Rmd
Worked on responses to R1/ improving results section

![**Figure `r fig_variable_mapping`. Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.** For each C pool, we show ForC variables corresponding to the stock, net annual increment, influx, and outflux. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter layer (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon. **This variable is techically EFDB-relevant but not selected for submission because their is no corresponding influx variable.](figures_tables/C_variable_mapping.png)

@@

As of `r format(Sys.time(), '%B %d, %Y')`<!--inserts date upon knit-->, ForC (v4.0) contained `r n_independent_records_ForC` independent records (`r n_records_ForC` total), `r n_ind_records_EFDB_variables` of which were for the `r n_relevant_variables` variables relevant to EFDB (Fig. `r fig_variable_mapping`).
We had reviewed or added `r n_records_reviewed_or_added` records for submission to EFDB, `r n_records_sent_to_EFDB` of which had been submitted, and `r n_records_posted_EFDB` of which had been reviewed by EFDB's review panel, accepted, and posted (`r table_variables`).
The `r round((n_records_reviewed_or_added -n_records_sent_to_EFDB )/n_records_reviewed_or_added,2)*100`% discrepancy between records we reviewed and those submitted to EFDB was due to the presence of digitized records and records where a variable's value had been calculated as the sum or difference of related variables rather than presented directly in the text.
The discrepancy between the number of records submitted and that posted to EFDB is because some records (`r round((1-(n_records_posted_EFDB/n_records_sent_to_EFDB)),2)*100`%) were deemed not applicable to EFDB by the review panel.
Reasons included -- but were not limited to -- non-applicability to the IPCC methodology of the variable submitted (e.g., net ecosystem CO~2~ exchange, litter - OL layer; subsequently excluded from lists of relevant variables and counts of relevant records), inadequate information on the quality of data and thus on its replicability (e.g., confidence interval/uncertainty), and vaguely described disturbance histories (e.g., "moderately"/ "severely" burned). 
As of `r format(Sys.time(), '%B %d, %Y')`, the `r n_records_sent_to_EFDB` ForC records posted in EFDB represented  ~`r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables,2)*100`% of the EFDB-relevant data in ForC and `r percent_forest_records`% of the total EFDB records for forest land.

ForC v4.0 contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables (or closely-related variable groups) relevant to EFDB (Fig. `r fig_variable_mapping`,  `r table_variables`).
Records were submitted for `r n_relevant_variables_records_sent` variables (or closely-related variable groups), including variables from each C pool (`r table_variables`).
The records available in ForC v4.0 -- and the subset submitted to EFDB -- were very unevenly distributed across variables (`r table_variables`). 
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records submitted were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round(n_records_sent_rootbiomass/n_records_sent_to_EFDB,2)*100`% for root biomass, 
`r round(n_records_sent_rootbiomass_components/n_records_sent_to_EFDB,2)*100`% for components of root biomass (coarse or fine roots), 
`r round(n_records_sent_deadwood/n_records_sent_to_EFDB,2)*100`% for dead wood, 
`r round(n_records_sent_deadwood_components/n_records_sent_to_EFDB,2)*100`% for components of dead wood (standing or fallen), 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (entire O horizon),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for SOM/ SOC.
Increment records totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records submitted, all but four of which described aboveground biomass.
The remaining `r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`% of records submitted described fluxes, all of which were either inputs or outputs to the aboveground biomass pool, a subset of which also described inputs to the dead wood or litter pool (Fig. `r fig_variable_mapping`,`r table_variables`).

ForC v4.0 records and the subset submitted to EFDB were distributed across all forested continents, biomes, and forest types, albeit unevenly (Figs. `r fig_map`-`r fig_histograms`).
Relative to forested area, submitted data under-represented the tropics and over-represented temperate regions, mirroring the composition of ForC (Figs. `r fig_map`, `r fig_histograms`b).
Submitted data covered all forested climates with substantial land area except the colder boreal regions (boreal tundra woodland; Figs. `r fig_map`, `r fig_climate`) and were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`).
Among the records submitted to EFDB, the largest number came from North America, followed by Asia, South America, and Africa, with strongly disproportionate representation of North America and under-representation of South America, Africa, Europe, Australia, and Oceania (Fig. `r fig_histograms`c).
In terms of FAO ecozones, boreal coniferous forest, temperate continental forest, and temperate mountain systems had the most records and were the best represented relative to their global area, whereas other ecozones were poorly represented relative to their global area (Fig. `r fig_histograms`b).
By far the most records and best representation relative to global area were for needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest number of records came from mature forests (>100 years), followed by young and intermediate-aged stands, but young (<20 year) stands were best represented relative to their global area (Fig. `r fig_histograms`d).

>>>>

![**Figure `r fig_variable_mapping`. Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.** For each C pool, we show ForC variables corresponding to the stock, net annual increment, influx, and outflux. Most, but not all, of the `r n_relevant_variables` EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter layer (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon. **This variable is techically EFDB-relevant but not selected for submission because their is no corresponding influx variable.](figures_tables/C_variable_mapping.png)

@@

As of `r format(Sys.time(), '%B %d, %Y')`<!--inserts date upon knit-->, we had submitted only a fraction (~`r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables,2)*100`%) of relevant ForC records to EFDB, but these already comprised a substantial portion (`r percent_forest_records`%) of EFDB records for forest land.
Specifically, of the `r n_records_ForC` records in ForC (v4.0), ~`r n_ind_records_EFDB_variables` were independent records of EFDB-relevant variables (Fig. `r fig_variable_mapping`,  `r table_variables`).
We reviewed `r n_records_checked` records for submission to EFDB, of which
`r round((n_records_checked -(n_records_sent_to_EFDB - n_records_added))/n_records_checked,2)*100`% were inappropriate, mostly because values were digitized from graphs or calculated from related variables rather than directly presented.
We also added `r n_records_added` new records, mostly because EFDB required that all relevant records in a publication be submitted together. 
Of `r n_records_sent_to_EFDB` submitted records, EFDB's review panel accepted and posted `r n_records_posted_EFDB` but deemed `r round((1-(n_records_posted_EFDB/n_records_sent_to_EFDB)),2)*100`% not applicable.
Reasons included -- but were not limited to -- non-applicability to the IPCC methodology of the variable submitted (e.g., net ecosystem CO~2~ exchange, litter - OL layer; subsequently excluded from lists of relevant variables and counts of relevant records), inadequate information on the quality of data and thus on its replicability (e.g., confidence interval/uncertainty), and imprecisely described disturbance histories (e.g., "moderately"/ "severely" burned).
An iterative process of data submission review by the EFDB Technical Support Unit and review panel greatly improved but did not perfect (100% acceptance rate) our process.

In terms of variables, ForC v4.0 contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables (or closely-related variable groups) relevant to EFDB (Fig. `r fig_variable_mapping`,  `r table_variables`).
The records submitted to EFDB (`r n_relevant_variables_records_sent` variables) were very unevenly distributed across variables, mirroring the composition of ForC (`r table_variables`). 
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records submitted were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round((n_records_sent_rootbiomass+n_records_sent_rootbiomass_components)/n_records_sent_to_EFDB,2)*100`% for root biomass or its components, 
`r round((n_records_sent_deadwood+n_records_sent_deadwood_components)/n_records_sent_to_EFDB,2)*100`% for dead wood or its components, 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (O horizon),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for soils.
Increment records (mostly for aboveground biomass) totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records submitted.
The remaining `r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`% of records submitted described fluxes, all of which were either inputs or outputs to the aboveground biomass pool, with some also describing inputs to dead wood or litter pools (Fig. `r fig_variable_mapping`,`r table_variables`).
Many of the EFDB-relevant variables remain poorly represented in ForC (`r table_variables`).

ForC v4.0 records -- and the subset submitted to EFDB -- were unevenly distributed across all forested continents, biomes, and forest types (Figs. `r fig_map`-`r fig_histograms`).
Relative to forested area, submitted data under-represented the tropics and over-represented temperate regions, mirroring the composition of ForC (Figs. `r fig_map`, `r fig_histograms`b).
Submitted data covered all forested climates with substantial land area except the colder boreal regions (boreal tundra woodland; Figs. `r fig_map`, `r fig_climate`) and were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`).
Among the records submitted to EFDB, the largest number came from North America, followed by Asia, South America, and Africa, with strongly disproportionate representation of North America and under-representation of South America, Africa, Europe, Australia, and Oceania (Fig. `r fig_histograms`c).
In terms of FAO ecozones, boreal coniferous forest, temperate continental forest, and temperate mountain systems had the most records and were the best represented relative to their global area, whereas other ecozones were poorly represented relative to their global area (Fig. `r fig_histograms`b).
By far the most records and best representation relative to global area were for needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest number of submitted records came from mature forests (>100 years), but young (<20 year) stands were best-represented relative to their global area, while 20-100 year stands were under-represented (Fig. `r fig_histograms`d).
ForC contains the records needed to provide more balanced geographical and forest type representation via selective data submission (Figs. `r fig_map`-`r fig_histograms`).
